---
title: "SimSpec_pyR_NDC"
author: "Arthur de Grandpré"
date: "19 juillet 2020"
output: html_document
---

```{r setup, include=FALSE}
rm(list=ls())
gc()

library(ncdf4)
library(raster)
library(sf)
library(raster)
library(rgdal)
library(rgeos)
library(tidyverse)
library(lubridate)
library(zoo)
library(parallel)
library(snowfall)
```

Nic's original workflow :  
Load S2A and S2B data separately, separate water and land pixels, apply simspec correction on a per pixel basis, based on an empirical ratio for each satellite, and calibrated to bands 865 and 783

What we have : 
from the NC files, land, water and vegetation are already separated, but the extraction got rid of the land pixels.

In order to match the workflows, we have to combine the simspec with our extraction and masking steps


```{r}
path_nc = "D:/nic_transfer/" # path to all .nc files
ncdir = dir(path_nc, pattern=".nc$")

path_poly = "./data/mask_nic_20191126/" # path to all ponds .shp files
polydir = dir(path_poly, pattern=".shp$")

# Assuming that all the .nc files will be structured in the same way, run this loop and voila :) 

nc_extract_simspec = function(x,ncdir,polydir){
# 
# x=1
# print(paste0("image ",i," out of ",length(ncdir)))
nc = nc_open(paste0(path_nc,ncdir[x]))
# nc1 # summary of the netcdf dataset

nc_proj4 = ncatt_get(nc,0, "proj4_string")$value
nc_res = ncatt_get(nc,0, "pixel_size")$value
nc_extent_x = ncatt_get(nc,0, "xrange")$value
nc_extent_y = ncatt_get(nc,0, "yrange")$value

var_names = attributes(nc$var)$names
bands = grep("^rhos",var_names, value=T)

r = raster()
for(b in 1:length(bands)){
  r = stack(r, raster(t(ncvar_get(nc, bands[b]))))
}

if(is.na(getValues(r[[1]])[1])==T){
  print(paste0("empty raster at ",x))
  next
}

r = brick(r)
extent(r) = c(nc_extent_x,nc_extent_y)
res(r) = nc_res
proj4string(r) = nc_proj4
names(r) = bands

r = projectRaster(r, crs="+init=epsg:4326")
output_df=data.frame()
for(j in 1:length(polydir)){
# j=1
  poly = st_read(paste0(path_poly,polydir[j]), quiet=T)
  
  cr = crop(r,poly)
  mr = mask(cr, poly)
  rmr = mask(r, poly, inverse=T)
  
  envs = c()
  for(b in 1:nlayers(rmr)){
  envs=append(envs, mean(getValues(rmr[[b]]),na.rm=T)) 
  }
  rm(rmr);rm(cr)
  
  landm865 = envs[9]
  landm783 = envs[7]
  
  dfmr = as.data.frame(mr, xy=T)
  dfmr = dfmr[complete.cases(dfmr),]
  dfmr$date = substr(ncdir[x],9,18)
  dfmr$satellite = substr(ncdir[x],1,3)
  dfmr$pond = gsub("_Polygon.shp","",polydir[j])
  
  class = apply(dfmr,1, function(x){
    if(x[4]>0.2){return("clouds")}else{
    if(x[10]>0.08){return("vegetation")}else{
    return("water")
  }}})
  
  # ggplot(dfmr, aes(x=class))+
  #   geom_bar()
  
  dfmr = cbind(dfmr,class)
  
  if("water" %in% unique(dfmr$class)){
    dfmrw = subset(dfmr, dfmr$class=="water")
    colnames(dfmrw) = colnames(dfmr)
    dfmrw$sol=NA
    
    for(k in 1:dim(dfmrw)[1]){
      tdf = dfmrw[k,]
      s2ac = 0.81936292024314039
      s2bc = 0.8296305320098123
      
      optif = function(x){
        tv = (tdf$rhos_865-x*landm865)/(tdf$rhos_783-x*landm783)
        if(tdf$satellite=="S2A"){
          return(abs(tv-s2ac))}else{
            return(abs(tv-s2bc))}
      }
      
      sol = optim(0,
                  optif,
                  lower = 0,
                  upper = 1)
      
      dfmrw[k,ncol(dfmrw)] = sol$par
    }
    
    dfc = dfmrw[,3:13]
    for(i2 in 1:ncol(dfc)){
      dfc[,i2]=dfc[,i2]-dfmrw[,18]*envs[i2]
    }
    
    dfc2 = dfmrw[,1:17]
    dfc2[,3:13]=dfc
    
    dfmr2 = subset(dfmr, dfmr$class=="vegetation")
    dfmr2 = rbind(dfmr2,dfc2)
    dfmr  = dfmr2
  }
  output_df=rbind(output_df,dfmr)
}
  colnames(output_df)=colnames(dfmr)
  write.csv(output_df, paste0("D:/Data/simspec2020/",x,"_nc_extract_simspec.csv"))
}


sfInit(parallel=TRUE, cpus=parallel::detectCores())
sfExport("ncdir","polydir","nc_extract_simspec","path_nc","path_poly")
sfLibrary(rgdal)
sfLibrary(ncdf4)
sfLibrary(raster)
sfLibrary(sp)
sfLibrary(sf)

out.lst = sfClusterApplyLB(1:length(ncdir),
                           function(x){nc_extract_simspec(x,ncdir,polydir)})

sfStop()

```

```{r}
csvs = dir("D:/Data/simspec2020/",pattern=".csv*",full.names=T)
# install.packages("vroom")
library(vroom)
df = vroom(csvs, col_names=F, skip=1)

dft = read.csv(csvs[1])
colnames(df)=colnames(dft)

write.csv(df, "./data/simspec2020.csv")
```


# VALIDATION

```{r}
library(leaflet)

path_nc = "D:/nic_transfer/" # path to all .nc files
ncdir = dir(path_nc, pattern=".nc$")

path_poly = "./data/mask_nic_20191126/" # path to all ponds .shp files
polydir = dir(path_poly, pattern=".shp$")

# Assuming that all the .nc files will be structured in the same way, run this loop and voila :) 

x=9

print(paste0("image ",x," out of ",length(ncdir)))
nc = nc_open(paste0(path_nc,ncdir[x]))
# nc1 # summary of the netcdf dataset

nc_proj4 = ncatt_get(nc,0, "proj4_string")$value
nc_res = ncatt_get(nc,0, "pixel_size")$value
nc_extent_x = ncatt_get(nc,0, "xrange")$value
nc_extent_y = ncatt_get(nc,0, "yrange")$value

var_names = attributes(nc$var)$names
bands = grep("^rhos",var_names, value=T)

r = raster()
for(b in 1:length(bands)){
  r = stack(r, raster(t(ncvar_get(nc, bands[b]))))
}

if(is.na(getValues(r[[1]])[1])==T){
  print(paste0("empty raster at ",x))
  next
}

r = brick(r)
extent(r) = c(nc_extent_x,nc_extent_y)
res(r) = nc_res
proj4string(r) = nc_proj4
names(r) = bands

r = projectRaster(r, crs="+init=epsg:4326")

j=1

poly1 = st_read(paste0(path_poly,polydir[j]), quiet=T)
poly2 = st_read(paste0(path_poly,polydir[j+1]), quiet=T)
poly3 = st_read(paste0(path_poly,polydir[j+2]), quiet=T)

leaflet() %>% 
  addTiles() %>% 
  addRasterImage(r[[4]]) %>% 
  addPolygons(data=poly1) %>% 
  addPolygons(data=poly2) %>% 
  addPolygons(data=poly3)

```

```{r}

df = read.csv("./data/simspec2020.csv")
df$date = ymd(df$date)

x1 = ymd(substr(sub(".*MSI_","",ncdir),start=1,stop=10)[x])

d1 = df %>%
  filter(df$date==x1) %>% 
  st_as_sf(coords = c("x","y"),
           crs = "+init=epsg:4326")

library(leaflet)

pal1 = colorNumeric(
  palette = "RdYlGn",
  domain = d1$rhos_833
)
# 
# leaflet(d1) %>%
#   addTiles() %>% 
#   addCircleMarkers(stroke = F,
#                    fillColor=~pal1(rhos_833),
#                    fillOpacity = 0.5) %>%
#   addLegend("bottomright",
#             pal=pal,
#             values=~rhos_833,
#             title= "NIR reflectance")


# Create a palette that maps factor levels to colors
pal2 <- colorFactor(c("white","red","navy"), domain = c("water","vegetation","clouds"))

leaflet(d1) %>%
  
  addTiles(
    group = "base OSM") %>%
  
  addCircleMarkers(
    color = ~pal2(class),
    stroke = FALSE, fillOpacity = 0.5,
    group = "classes") %>% 
  
  addLegend(
    "bottomright",
    pal = pal2,
    values = ~class,
    title = "classes",
    group = "classes") %>% 
  
  addCircleMarkers(
    stroke = F,
    fillColor = ~pal1(rhos_833),
    fillOpacity = 0.5,
    group = "NIR") %>% 
  
  addLegend(
    "bottomright",
    pal = pal1,
    values = ~rhos_833,
    title = "NIR reflectance",
    group = "NIR") %>% 
  
  addRasterImage(
    r[[8]],
    group = "base NIR") %>% 
  
  addLayersControl(
    baseGroups = c("base OSM"),
    overlayGroups = c("classes", "NIR", "base NIR"),
    options = layersControlOptions(collapsed = FALSE))
  
```

