---
title: "water_veg_mask"
author: "Arthur de Grandpré"
date: "17 novembre 2019"
output: html_document
---

```{r setup}

rm(list=ls())

library(raster)
library(rgdal)
library(rgeos)
library(sf)
library(leaflet)
library(mapview)

```

```{r load images}

# where to find the images
s2_ac_dir = "data/heavy"
s2_ac_imgs = dir(s2_ac_dir, pattern = ".tif$", recursive=T)
s2_ac_img = paste0(s2_ac_dir,"/",s2_ac_imgs[1])

# the images CRS according to SNAP
WGS84_31N = "+init=epsg:32631"

# loading the raster
r = brick(s2_ac_img)
r
# setting the right layer names
layers = c("x",
           "y",
           "443",
           "492",
           "560",
           "665",
           "704",
           "740",
           "783",
           "833",
           "865",
           "1614",
           "2202")
names(r) = layers
r = dropLayer(r, c("x","y"))
r = r/255
plot(r)

```

```{r subset image}

#mapRGB = viewRGB(r, 4,3,2, quantiles = c(0.05,0.95))
#subset = drawFeatures(map = mapRGB)
#st_write(subset, "data/pond_area.gpkg", layer="sub_extent", driver ="GPKG")
subset = st_read("data/pond_area.gpkg")
rs = crop(r, extent(subset))

```

# Pond masking validation (extra)

Since the masking seems to perform very badly, let's look at the polygon masks used to subset the ponds from the rasters

```{r}
library(rgdal)
library(raster)
rm = trim(raster(paste0("./data/mask_nic_20191126/",dir("./data/mask_nic_20191126/", pattern=".tif$")[1])))
rm2 = rasterToPolygons(rm,dissolve=T)
rm3 = spTransform(rm2, CRS("+init=epsg:4326"))
plot(rm3)

rm4 = readOGR(paste0("./data/mask_nic_20191126/",dir("./data/mask_nic_20191126/", pattern=".shp$")[1]))
```


```{r visualise image}

library(mapedit)
library(leaflet)

val = as.numeric(c(0:1))
pal = colorNumeric("RdYlGn",val)

map = leaflet() %>%
  addTiles() %>% 
  addRasterImage(rs$X833,
                 colors = pal,
                 opacity = 0.5,
                 group="B8") %>% 
  addPolygons(data=rm4, group="tif") %>% 
  addPolygons(data=rm3, group="shp") %>% 
  addLayersControl(overlayGroups=c("B8","tif","shp"))
map

library(htmlwidgets)
saveWidget(map, "fit_test_pond.html", selfcontained = T)

options(mapviewMaxPixels = 1000000)

comp = viewRGB(rs, 4,3,2, quantiles = c(0.01, 0.9))
comp

```

there is a big gap between the polygons and the actual ponds. it seems to be from the .nc file, maybe when it is transformed into .tiff  
   
let's try to read the netcdf file directly
```{r eval=FALSE, include=FALSE}
#install.packages("RNetCDF")
rm(list=ls())
path = "./data/heavy/wetransfer-e8e5fa/"
ncdir = dir(path, pattern=".nc$")
 library(RNetCDF)
 nc1 = read.nc(paste0(path,ncdir[1]))
### seems like it's not a true netcdf file, probably hdf5

# install.packages("BiocManager")
# BiocManager::install("rhdf5")

library(rhdf5)

h5_index = h5ls(paste0(path,ncdir[1]), all=T)

h5_index$name
dfrnc1 = cbind(as.vector(h5read(paste0(path,ncdir[1]), name=h5_index$name[3])),
               as.vector(h5read(paste0(path,ncdir[1]), name=h5_index$name[2])),
               as.vector(h5read(paste0(path,ncdir[1]), name=h5_index$name[6])),
               as.vector(h5read(paste0(path,ncdir[1]), name=h5_index$name[7])),
               as.vector(h5read(paste0(path,ncdir[1]), name=h5_index$name[8])),
               as.vector(h5read(paste0(path,ncdir[1]), name=h5_index$name[9])),
               as.vector(h5read(paste0(path,ncdir[1]), name=h5_index$name[10])),
               as.vector(h5read(paste0(path,ncdir[1]), name=h5_index$name[11])),
               as.vector(h5read(paste0(path,ncdir[1]), name=h5_index$name[12])),
               as.vector(h5read(paste0(path,ncdir[1]), name=h5_index$name[13])),
               as.vector(h5read(paste0(path,ncdir[1]), name=h5_index$name[14])),
               as.vector(h5read(paste0(path,ncdir[1]), name=h5_index$name[4])),
               as.vector(h5read(paste0(path,ncdir[1]), name=h5_index$name[5])))

head(dfrnc1)
write.csv(dfrnc1, "./data/heavy/nc1.csv")

library(raster)
rasterOptions(tmpdir="D:/Rtemp")
rasterOptions(progress="window")
rasterOptions(timer=T)
rasterOptions(todisk=T)
# writeRaster(rasterFromXYZ(read.csv("./data/heavy/nc1.csv")), filename(".data/heavy/nc1.tif"),format="GTiff") This solution is way too long

lonr = raster(t(h5read(paste0(path,ncdir[1]), name=h5_index$name[3])))
latr = raster(t(h5read(paste0(path,ncdir[1]), name=h5_index$name[2])))

rt1 = raster(t(h5read(paste0(path,ncdir[1]), name=h5_index$name[13])))


rsnap = rt1

extent(rt1) = c(lonr@data@min,lonr@data@max,latr@data@min,latr@data@max)
res(rt1) = c((lonr@data@max-lonr@data@min)/lonr@ncols,(latr@data@max-latr@data@min)/lonr@nrows)
crs(rt1) = CRS("+init=epsg:4326")
library(leaflet)
library(sp)
library(rgdal)

rm4 = readOGR(paste0("./data/mask_nic_20191126/",dir("./data/mask_nic_20191126/", pattern=".shp$")[1]))

tif_r = raster(paste0("./data/crop_rasters/",dir("./data/crop_rasters/")[1]), band=8)
tif_r

extent(rsnap) = c(654900,675540,5644440,5656200)
res(rsnap) =c(10,10)
crs(rsnap) = CRS("+proj=utm +zone=31 +datum=WGS84 +units=m +no_defs")
rsnap

rsnapg = projectRaster(rsnap, crs="+init=epsg:4326")

map = leaflet() %>% 
  addTiles() %>% 
  addRasterImage(rt1,
                 opacity=1,
                 group = "raster_nc") %>%
  addRasterImage(tif_r,
                 opacity=1,
                 group = "raster_tif") %>%
  addRasterImage(rsnapg,
                 opacity=1,
                 group = "snap_manual") %>% 
  addPolygons(data=rm4,
              group = "polygon") %>% 
  addLayersControl("topleft", overlayGroups = c("polygon","raster_nc","raster_tif","snap_manual")) ; map
library(htmlwidgets)
saveWidget(map,"map_tif_nc_snap_polygon.html",selfcontained=T)
rt1
rasterOptions(todisk=F)
h5closeAll()
```

The last chunck is very chaotic because of multiple attemps, but the data is in fact a WGS84 UTM 31N grid and the metadata is available in the SNAP software, but not visible in R ... using the contained lat lon data doesnt give a proper geolocation, nor the SNAP exported tif files. Thus, we need to access the metadata from within R and manually project this data. Let's try to do this in a clean way.

```{r}
library(ncdf4)
rm(list=ls())
path = "./data/heavy/wetransfer-e8e5fa/"
ncdir = dir(path, pattern=".nc$")

nc = nc_open(paste0(path,ncdir[1]))
# nc1 # summary of the netcdf dataset

nc_proj4 = ncatt_get(nc,0, "proj4_string")$value
nc_res = ncatt_get(nc,0, "pixel_size")$value
nc_extent_x = ncatt_get(nc,0, "xrange")$value
nc_extent_y = ncatt_get(nc,0, "yrange")$value

var_names = attributes(nc$var)$names
bands = grep("^rhos",var_names, value=T)

r = raster()
for(i in 1:length(bands)){
r = stack(r, raster(t(ncvar_get(nc, bands[i]))))
}
r = brick(r)

extent(r) = c(nc_extent_x,nc_extent_y)
res(r) = nc_res
proj4string(r) = nc_proj4
names(r) = bands

rp = projectRaster(r, crs="+init=epsg:4326")
writeRaster(rp, paste0("./data/heavy/"))
#  leaflet() %>% 
#    addTiles() %>% 
#    addRasterImage(rp$rhos_833)

```


```{r polygon library drawing}

# some_clouds = mapedit::drawFeatures(map = comp)
# st_write(some_clouds, "data/training_features.gpkg", layer="clouds", driver = "GPKG", delete_layer = T)

# some_ponds = mapedit::drawFeatures(map = comp)
# st_write(some_ponds, "data/training_features.gpkg", layer="ponds", driver ="GPKG", delete_layer = T)

# some_urban = mapedit::drawFeatures(map = comp)
# st_write(some_urban, "data/training_features.gpkg", layer="urban", driver ="GPKG", delete_layer = T)

# some_forest = mapedit::drawFeatures(map = comp)
# st_write(some_forest, "data/training_features.gpkg", layer="forest", driver ="GPKG")

# some_fields = mapedit::drawFeatures(map = comp)
# st_write(some_fields, "data/training_features.gpkg", layer="fields", driver ="GPKG")

# some_water = mapedit::drawFeatures(map = comp)
# st_write(some_water, "data/training_features.gpkg", layer="water", driver ="GPKG")

 # some_veg = mapedit::drawFeatures(map = comp)
 # st_write(some_veg, "data/training_features.gpkg", layer="veg", driver ="GPKG")

```

```{r polygon library reading}

st_layers("data/training_features.gpkg")

veg = st_read("data/training_features.gpkg", "veg")
water = st_read("data/training_features.gpkg", "water")

comp_w = addFeatures(comp, data=water)
comp_wv = addFeatures(comp_w, data=veg)
comp_wv

```

```{r single image value comparison}

water_r = mask(rs, water)
veg_r = mask(rs, veg)

dev.off()
boxplot(veg_r, main = "veg", col = "green", notch = T)
boxplot(water_r, main = "water", col = "blue", add = T, notch = T)

hist(veg_r)
hist(water_r)

```

```{r multiple image fit}

for(i in 1:length(s2_ac_imgs)) {
s2_ac_img = paste0(s2_ac_dir,"/",s2_ac_imgs[i])

# loading the raster
r = brick(s2_ac_img)

# setting the right layer names
layers = c("x",
           "y",
           "443",
           "492",
           "560",
           "665",
           "704",
           "740",
           "783",
           "833",
           "865",
           "1614",
           "2202")
names(r) = layers
r = dropLayer(r, c("x","y"))
r = r/255

subset = st_read("data/pond_area.gpkg")
rs = crop(r, extent(subset))

writeRaster(rs, paste0("data/crop_rasters/",gsub(".*fa/","",s2_ac_img)), format="GTiff")    
}

```

```{r multiple raster view}
crops = dir("data/crop_rasters/", patter=".tif$")

val = c(0,1)
pal = colorNumeric("RdYlGn",
                   val,
                   na.color = "transparent")

leaflet() %>% 
  
  addRasterImage(raster(paste0("data/crop_rasters/",crops[1]),band=4),
                 col = pal, group = "r1") %>% 
  
  addRasterImage(raster(paste0("data/crop_rasters/",crops[2]),band=4),
                 col = pal, group = "r2") %>% 
  
  addRasterImage(raster(paste0("data/crop_rasters/",crops[3]),band=4),
                 col = pal, group = "r3") %>% 
  
  addRasterImage(raster(paste0("data/crop_rasters/",crops[4]),band=4),
                 col = pal, group = "r4") %>% 
  
  addPolygons(data = water,
              color = "blue" ,
              group = "water") %>% 

  addPolygons(data = veg,
              color = "green",
              group = "veg"  ) %>% 
  
  addLayersControl("topright",
                   overlayGroups = c("r1","r2","r3","r4","water","veg"))

```

Alignement seems to be perfect, let's see how the scenes compare in the RGB view

```{r}
m1 = viewRGB(brick(paste0("data/crop_rasters/",crops[1]), r=4,g=3,b=2, quantiles=NULL))
m2 = viewRGB(brick(paste0("data/crop_rasters/",crops[2]), r=4,g=3,b=2, quantiles=NULL))
m3 = viewRGB(brick(paste0("data/crop_rasters/",crops[3]), r=4,g=3,b=2, quantiles=NULL))
m4 = viewRGB(brick(paste0("data/crop_rasters/",crops[4]), r=4,g=3,b=2, quantiles=NULL))
mA = sync(m1,m2,m3,m4, ncol=2)
mA

```

```{r}
library(htmlwidgets)
library(htmltools)
save_tags <- function (tags, file, selfcontained = F, libdir = "./lib") 
{
  if (is.null(libdir)) {
    libdir <- paste(tools::file_path_sans_ext(basename(file)), 
                    "_files", sep = "")
  }
  htmltools::save_html(tags, file = file, libdir = libdir)
  if (selfcontained) {
    if (!htmlwidgets:::pandoc_available()) {
      stop("Saving a widget with selfcontained = TRUE requires pandoc. For details see:\n", 
           "https://github.com/rstudio/rmarkdown/blob/master/PANDOC.md")
    }
    htmlwidgets:::pandoc_self_contained_html(file, file)
    unlink(libdir, recursive = TRUE)
  }
  return(file)
}
save_tags(mA, "nic_4pane_432.html", selfcontained = T)
```

Since the ground reflectance values seem very different for every images, we will tend towards the use of band ratios instead of single bands for classification.

```{r}
layers2 = layers[3:13]

r1 = brick(paste0("data/crop_rasters/",crops[1])) ; names(r1) = layers2
r2 = brick(paste0("data/crop_rasters/",crops[2])) ; names(r2) = layers2
r3 = brick(paste0("data/crop_rasters/",crops[3])) ; names(r3) = layers2
r4 = brick(paste0("data/crop_rasters/",crops[4])) ; names(r4) = layers2

r1$NDVI = (r1$X833-r1$X665) / (r1$X833+r1$X665)
r2$NDVI = (r2$X833-r2$X665) / (r2$X833+r2$X665)
r3$NDVI = (r3$X833-r3$X665) / (r3$X833+r3$X665)
r4$NDVI = (r4$X833-r4$X665) / (r4$X833+r4$X665)

```


```{r}
water1 = mask(r1, water)
water2 = mask(r2, water)
water3 = mask(r3, water)
water4 = mask(r4, water)

veg1 = mask(r1, veg)
veg2 = mask(r2, veg)
veg3 = mask(r3, veg)
veg4 = mask(r4, veg)

```

```{r}
dev.off()
par(mfrow=c(2,2))
boxplot(water1)
boxplot(water2)
boxplot(water3)
boxplot(water4)
boxplot(veg1)
boxplot(veg2)
boxplot(veg3)
boxplot(veg4)
```

an NDVI treshold of about 0 seems to work fine. let'S look at the maps.

```{r}
r_ndvi = brick(r1$NDVI,r2$NDVI,r3$NDVI,r4$NDVI)

val = c(-1.1,1.1)
pal = colorNumeric("RdYlGn",
                   val,
                   na.color = "transparent")


leaflet() %>% 
  
  addRasterImage(r_ndvi$NDVI.1,
                 col = pal, group = "r1",
                 layerId = "r1") %>% 
  
  addRasterImage(r_ndvi$NDVI.2,
                 col = pal, group = "r2",
                 layerId = "r2") %>% 
  
  addRasterImage(r_ndvi$NDVI.3,
                 col = pal, group = "r3",
                 layerId = "r3") %>% 
  
  addRasterImage(r_ndvi$NDVI.4,
                 col = pal, group = "r4",
                 layerId = "r4") %>% 
  
  addPolygons(data = water,
              color = "blue" ,
              group = "water") %>% 

  addPolygons(data = veg,
              color = "green",
              group = "veg"  ) %>% 
  
  addLegend("bottomright",
            pal = pal,
            values = val,
            title = "NDVI") %>% 
  
  addMouseCoordinates() %>% 
  
  addImageQuery(r_ndvi$NDVI.1, type="mousemove", layerId = c("r1")) %>% 
  addImageQuery(r_ndvi$NDVI.2, type="mousemove", layerId = c("r2")) %>% 
  
  
  addLayersControl("topright",
                   overlayGroups = c("r1","r2","r3","r4","water","veg"))
```

```{r}
r_ndvi = brick(r1$NDVI,r2$NDVI,r3$NDVI,r4$NDVI)

val = c(-1.1,1.1)
pal = colorNumeric("RdYlGn",
                   val,
                   na.color = "transparent")

ndvi1 = leaflet() %>% 
  addRasterImage(r_ndvi$NDVI.1,
                 col = pal,
                 layerId = "ndvi") %>% 
  addPolygons(data = water,
              color = "blue" ,
              group = "water") %>% 
  addPolygons(data = veg,
              color = "green",
              group = "veg"  ) %>% 
  addLegend("bottomright",
            pal = pal,
            values = val,
            title = "NDVI") %>% 
  addMouseCoordinates() %>% 
  addImageQuery(r_ndvi$NDVI.1, type="mousemove", layerId = c("ndvi"))

ndvi2 = leaflet() %>% 
  addRasterImage(r_ndvi$NDVI.2,
                 col = pal,
                 layerId = "ndvi") %>% 
  addPolygons(data = water,
              color = "blue" ,
              group = "water") %>% 
  addPolygons(data = veg,
              color = "green",
              group = "veg"  ) %>% 
  addLegend("bottomright",
            pal = pal,
            values = val,
            title = "NDVI") %>% 
  addMouseCoordinates() %>% 
  addImageQuery(r_ndvi$NDVI.2, type="mousemove", layerId = c("ndvi"))

ndvi3 = leaflet() %>% 
  addRasterImage(r_ndvi$NDVI.3,
                 col = pal,
                 layerId = "ndvi") %>% 
  addPolygons(data = water,
              color = "blue" ,
              group = "water") %>% 
  addPolygons(data = veg,
              color = "green",
              group = "veg"  ) %>% 
  addLegend("bottomright",
            pal = pal,
            values = val,
            title = "NDVI") %>% 
  addMouseCoordinates() %>% 
  addImageQuery(r_ndvi$NDVI.3, type="mousemove", layerId = c("ndvi"))

ndvi4 = leaflet() %>% 
  addRasterImage(r_ndvi$NDVI.4,
                 col = pal,
                 layerId = "ndvi") %>% 
  addPolygons(data = water,
              color = "blue" ,
              group = "water") %>% 
  addPolygons(data = veg,
              color = "green",
              group = "veg"  ) %>% 
  addLegend("bottomright",
            pal = pal,
            values = val,
            title = "NDVI") %>% 
  addMouseCoordinates() %>% 
  addImageQuery(r_ndvi$NDVI.4, type="mousemove", layerId = c("ndvi"))

ndvi_maps = sync(ndvi1,ndvi2,ndvi3,ndvi4)

save_tags <- function (tags, file, selfcontained = F, libdir = "./lib") 
{
  if (is.null(libdir)) {
    libdir <- paste(tools::file_path_sans_ext(basename(file)), 
                    "_files", sep = "")
  }
  htmltools::save_html(tags, file = file, libdir = libdir)
  if (selfcontained) {
    if (!htmlwidgets:::pandoc_available()) {
      stop("Saving a widget with selfcontained = TRUE requires pandoc. For details see:\n", 
           "https://github.com/rstudio/rmarkdown/blob/master/PANDOC.md")
    }
    htmlwidgets:::pandoc_self_contained_html(file, file)
    unlink(libdir, recursive = TRUE)
  }
  return(file)
}
save_tags(ndvi_maps, "nic_4pane_ndvi.html", selfcontained = T)
```

it seems like a reasonnable band ratio treshold would be ndvi = 0. But still, it should not be expected to hold up in every single cases

CLOUDS : 
```{r}
r_cloud = brick(r1$X492,r2$X492,r3$X492,r4$X492)

val = c(0,1)
pal = colorNumeric("RdYlGn",
                   val,
                   na.color = "transparent")

cloud1 = leaflet() %>% 
  addRasterImage(r_cloud$X492.1,
                 col = pal,
                 layerId = "B2") %>% 
  addPolygons(data = water,
              color = "blue" ,
              group = "water") %>% 
  addPolygons(data = veg,
              color = "green",
              group = "veg"  ) %>% 
  addLegend("bottomright",
            pal = pal,
            values = val,
            title = "B2") %>% 
  addMouseCoordinates() %>% 
  addImageQuery(r_cloud$X492.1, type="mousemove", layerId = c("cloud"))

cloud2 = leaflet() %>% 
  addRasterImage(r_cloud$X492.2,
                 col = pal,
                 layerId = "B2") %>% 
  addPolygons(data = water,
              color = "blue" ,
              group = "water") %>% 
  addPolygons(data = veg,
              color = "green",
              group = "veg"  ) %>% 
  addLegend("bottomright",
            pal = pal,
            values = val,
            title = "B2") %>% 
  addMouseCoordinates() %>% 
  addImageQuery(r_cloud$X492.2, type="mousemove", layerId = c("cloud"))

cloud3 = leaflet() %>% 
  addRasterImage(r_cloud$X492.3,
                 col = pal,
                 layerId = "B2") %>% 
  addPolygons(data = water,
              color = "blue" ,
              group = "water") %>% 
  addPolygons(data = veg,
              color = "green",
              group = "veg"  ) %>% 
  addLegend("bottomright",
            pal = pal,
            values = val,
            title = "B2") %>% 
  addMouseCoordinates() %>% 
  addImageQuery(r_cloud$X492.3, type="mousemove", layerId = c("cloud"))

cloud4 = leaflet() %>% 
  addRasterImage(r_cloud$X492.4,
                 col = pal,
                 layerId = "B2") %>% 
  addPolygons(data = water,
              color = "blue" ,
              group = "water") %>% 
  addPolygons(data = veg,
              color = "green",
              group = "veg"  ) %>% 
  addLegend("bottomright",
            pal = pal,
            values = val,
            title = "B2") %>% 
  addMouseCoordinates() %>% 
  addImageQuery(r_cloud$X492.4, type="mousemove", layerId = c("cloud"))

cloud_maps = sync(cloud1,cloud2,cloud3,cloud4)

save_tags <- function (tags, file, selfcontained = F, libdir = "./lib") 
{
  if (is.null(libdir)) {
    libdir <- paste(tools::file_path_sans_ext(basename(file)), 
                    "_files", sep = "")
  }
  htmltools::save_html(tags, file = file, libdir = libdir)
  if (selfcontained) {
    if (!htmlwidgets:::pandoc_available()) {
      stop("Saving a widget with selfcontained = TRUE requires pandoc. For details see:\n", 
           "https://github.com/rstudio/rmarkdown/blob/master/PANDOC.md")
    }
    htmlwidgets:::pandoc_self_contained_html(file, file)
    unlink(libdir, recursive = TRUE)
  }
  return(file)
}
save_tags(cloud_maps, "nic_4pane_clouds.html", selfcontained = T)
```

the cloud masking doesn't really have a suitable band ratios, so we have to use a single band treshold. those are expected to create a lot of noise around the masks, since a single value doesn't really fit all the data. b2's baseline is much higher in image 3 (winter), so it might behave poorly in some extreme cases, but a value of somewhere around 0.3 seems reasonable to limit data loss.



